---
title: "Plotting zipcode areas in R"
output:
  html_document: default
  html_notebook: default
---

Here's a quick and easy function to plot a shaded area of a zipcode over an open sourced map.

To do this, I first downloaded shapefiles of zipcode polygons from the [Census Bureau](https://www.census.gov/geo/maps-data/data/cbf/cbf_zcta.html) which I then read in using `rgdal::readOGR`.

```{r,message=FALSE,results='hide'}
library(rgdal)
zip <- readOGR(dsn = "census_zip_shpfiles", layer = "census_zip")
```

I then create a `get_zip` function which extracts the relevant polygon related to the `zipcode` variable. Note that this simply involves slicing the spatial polygon object using the ZCTA5CE10 variable in the `@data` slot.

```{r}
get_zip = function(zipcode) {
  return(zip[zip$ZCTA5CE10==zipcode,])
}
```

I then define a function `plot_zip` to shade the relevant zipcode on a map, using the Leaflet function.

```{r,message=FALSE}
library(leaflet); library(dplyr)
plot_zip = function(zipcode) {
  plt <- get_zip(zipcode) %>% leaflet() %>%
  addPolygons(stroke = TRUE,opacity = 1,fillOpacity = 0.5, smoothFactor = 0.5,color="black",weight = 0.5) %>%
  addTiles() 
  return(plt)
}
```

We can now use this function to easily take a quick look at various zipcodes. For example, I used to live in Ithaca, New york:

```{r ithaca, message=FALSE,results='hide'}
plot_zip(14850)
```

<center>
```{r, ref.label='ithaca',echo=FALSE}
```
</center>

<br>

Now let's take a look at my current zipcode
```{r logan,message=FALSE,results='hide'}
plot_zip(20005)
```

<center>
```{r,ref.label='logan', message=FALSE,echo=FALSE}
```
</center>

<br>

Finally, let's throw down a couple markers to point out a couple places I like to hang out on Friday after work. To do this, let's geocode a couple of addresses using `ggmap`.

```{r , message=FALSE,results='hide'}
x <- c("1430 Rhode Island Ave NW, Washington, DC 20005",
       "1333 14th St NW, Washington, DC 20005")

library(ggmap)
loc <- geocode(x)
```

The `loc` is a dataframe containing the lattitude and longitude of each of the adresses. We can then plot these points on the map we created earlier using the `addMarkers` Leaflet method.

```{r locations,results='hide'}
plot_zip(20005) %>% addMarkers(lng=loc$lon,lat=loc$lat)
```


<center>
```{r,ref.label='locations', echo=FALSE}
```
</center>

The code and census data can be downloaded from this [GitHub Repository](https://github.com/hautahi/ZipCodeMapR).